I owe my deepest gratitude to my beloved family, whose
boundless love and continuous encouragement was unremitting source of inspiration
for this work, specially my sister Afsoon who helped me a lot by all means in the past couple of years.  No words can express how grateful I am for your love and support and
how much I love and appreciate you.

My deepest gratitude is to my supervisor, Dr. Wahab Hamou-Lhadj. I have been amazingly fortunate to have a supervisor who gave me the freedom to explore on my own, and at the same time the guidance to recover when my steps faltered. His patience and support helped me overcome many crisis situations and finish this thesis. I am grateful to him for holding me to a high research standard and enforcing strict validations for each research result.

Beside I owe a dept of gratitute to my supervisor's postdoctoral fellow, Dr. Wael Khreich, who has been always there to listen and give advice. I am deeply grateful to him for the long discussions that helped me sort out the technical details of my work. 

Many friends have helped me stay sane through these difficult years. I greatly value their friendship and I deeply appreciate their belief in me. Special thanks to Ali Sehizadeh, not only for helping me through my master, but also for being warm hearted friends. I should have my special thank to my friends whose friendship going back to my bachelor days, thank you to the supportive friend Afshin Moazami.
Special thanks to Elnaz Davoodi and Nassim Masoudifar, for encouraging me along the way.
I am also indebted to my friends, Yashar Zeighami, Masoud Karzand, Pouya jabbari and Majid Laali for their support, comments and feedback.

